\vspace*{-8mm}\textbf{\textsc{Novelty Detection for Semantic Place Categorization}}

### 2011-06-22 thesis/summary-en.tex
Changes on the summary

\section{Problem and Goals}
This thesis focus then on the particular problem of identifying novel semantic categories for places
using
This thesis studies a proposed %~\cite{pronobis2011semmap}
process that utilizes spatial and semantic knowledge for semantic mapping using a
probabilistic graphical model,
and develops methods to perform detection of room categories not previously
known to the robot using that graphical model.

\section{Semantic Mapping}
Semantic mapping is the process by which an agent builds a representation of the environment and
enhances the identified entities with semantic values (i.e.\ human semantics).
In this thesis the semantic mapping process proposed by \cite{pronobis2011phd} was used as based.

\subsection{Spatial Knowledge Representation}
In order to perform semantic mapping an agent needs to possess knowledge on how to segment, identify
entities existent in the space by analysing the sensed properties. For that Pronobis~\cite{pronobis2011semmap} presents a
layered knowledge structure:
the sensory layer maintains an accurate and exact description of environment in a machine-friendly representation,
the place layer employs knowledge on how to segment and identify areas such as rooms,
the categorical layer defines knowledge on properties and categories that allow to extract
higher-level information from the sensed data
and the conceptual layer maintains knowledge on entities and their relations between them as well
information on how to map those to human semantics.

\subsection{Conceptual Map}
The semantic mapping is then based on the instantiation of a conceptual map that uses all the
ontology defined on the conceptual layer in order to build a probabilistic graphical model that
relates and propagates probabilities across all the entities detect and properties detected by the
other layers.

By using probabilistic graphical models, the conceptual map becomes then a generative model that
allows several inferences on variable configurations to be performed and for that serves as a useful
tool for novelty detection of unknown categories of some of its variables.

@@

\subsection{Assumption on a constant $P(novel)$}
The ratio introduced on the previous section allows to implement an optimal detector.
Though the factors involved need to be defined according to available data.

$P(\overline{novel})$ was considered constant through all possible situations. Although this
sounds a weak assumption note that in the case hat uses the ratio

\subsection{Modeling Conditional Probability}
Since the graphical model produced by the conceptual map tries to model the distribution of
variables assuming the knowledge of the agent holds true, it represents a method to model
$P(x|\overline{novel})$ since it allows to calculate the density probability that the set of features $x$ is
sensed given that all the variables and graph structure, including the category of
a specific room $a$ are correctly modelled by the agent knowledge.
This way the distribution modelled by a factor graph equivalent to the chain-graph used by
the conceptual map can be used as an approximation for the conditional probability of $x$.

\subsection{Modeling Unconditional Probability}
With no knowledge on the real distribution of all samples, the correct approach is to model it
with a uniform distribution. Nonetheless in the presented case, it was assumed that only the room
category of a specific room is unknown and that the remaining knowledge holds trues.
In that case, the correct approach is to replace the variable $a$, that represents the room where to
detect novelty, with a single factor connecting all variables directly dependent on it.

Due to its high-dimensionality, approximating that factor with unlabelled data is infeasible and
approximations have to be done. In this thesis work two approximations where tried: an uniform model
and an independent model as illustrated in \autoref{fig:models}. In the independent model, the
single\hyp{}connected factors are trained from unlabelled data and its expected that they create a better
approximation to the unconditional probability than an assumption on the uniform model.

\section{Results}
A synthetic dataset modelling the expected features when implementing on the real case and enforcing
a constant $P(novel)$ was used for testing the developed models.
The two suggested models where compared between them and to an optimal case across several situations
showing they behaviour as more information is sensed from the environment.

>>>>

\section{Problem and Goals}
The semantic mapping process plays an important role for allowing agents to reason at an high-level
and is totally defined by the knowledge the agent has on the spatial concepts and semantic categories.
For that reason there is an high interest in creating methods that are able to alert the agent for
gaps on its knowledge.
And with that purpose, this thesis, studies novelty detection techniques and how to apply them to
detect novel room categories during a semantic mapping process.
To this end, the semantic mapping process proposed by \cite{pronobis2011phd} is studied and a method
to detect novel categories using graphical models is suggested.

\section{Semantic Mapping}
Semantic mapping is the process by which an agent builds a representation of the environment and
enhances the identified entities with semantic values (i.e.\ human semantics).
In this thesis the semantic mapping process proposed by \cite{pronobis2011phd} is used as base.

\subsection{Spatial Knowledge Representation}
In order to perform semantic mapping an agent needs to possess knowledge on how to segment and
identify entities existent in space by analysing the sensed data. For that
Pronobis~\cite{pronobis2011semmap} presents a layered knowledge structure:
the \emph{sensory layer} maintains an accurate and exact description of environment in a machine-friendly representation,
the \emph{place layer} employs knowledge on how to segment and identify areas such as rooms,
the \emph{categorical layer} defines knowledge on properties and categories that allow to extract
higher-level information from the sensed data
and the \emph{conceptual layer} maintains knowledge on entities concepts, the relations between them
and information on how to map them to human semantics.

\subsection{Conceptual Map}
The semantic mapping is then based on the instantiation of a conceptual map that uses all the
ontology defined on the conceptual layer in order to build a probabilistic graphical model that
relates and propagates probabilities across all the entities and properties detected by the
other layers.

By using probabilistic graphical models, the conceptual map becomes then a generative model that
allows to perform inference on variable configurations. For that it is used as a base tool
for novelty detection of unknown categories on some of its variables.

@@

\subsection{Assumption on a constant $P(novel)$}
By considering $P(novel)$ to be constant a ratio between conditional and unconditional probabilities
of the sensed variables is obtained.
Although it sounds reasonable to assume the probability of specific room belong to a novel
category is constant, this thesis points that this assumption might be very strong and not
reasonable in a realistic scenario. The main point being that in this case the graph structure
changes between samples and some might be more prone to produce novel samples, as opposed to normal
classification scenarios where its clear that the category is indeed the only generator of features.

\subsection{Modeling Conditional Probability}
Since the graphical model produced by the conceptual map tries to model the distribution of
variables assuming the knowledge of the agent holds true, it represents a method to model
$P(x|\overline{novel})$.
%since it allows to calculate the density probability that the set of features $x$ is
%sensed given that all the variables and graph structure, including the category of
%a specific room $a$ are correctly modelled by the agent knowledge.
%This way the distribution modelled by a factor graph equivalent to the chain-graph used by
%the conceptual map can be used as an approximation for the conditional probability of $x$.

\subsection{Modeling Unconditional Probability}
With no knowledge on the real distribution samples, the correct approach is to model it
with a uniform distribution. Nonetheless in the presented case, it was assumed that only the room
category of a specific room is unknown and that the remaining knowledge holds trues.
In that case, the correct approach is to replace the variable that represents the room where to
detect novelty with a single factor connecting all variables directly dependent on it.

Due to its high-dimensionality, approximating that factor with unlabelled data is infeasible and
approximations have to be done. In this thesis two approximations were presented: an uniform model
and an independent model. In the independent model, the
single\hyp{}connected factors are trained from unlabelled data and its expected that they create a better
approximation to the unconditional probability than an assumption on the uniform model.

\section{Results}
A synthetic dataset modelling the expected features when implementing on the real case and enforcing
a constant $P(novel)$ was used for testing the developed models.
The two suggested models where compared between them and to an optimal case.
The tests across several situations shows how they behaviour change as more information is sensed
from the environment.
